---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# carfollowingmodels

<!-- badges: start -->
<!-- badges: end -->

The goal of `carfollowingmodels` is to make several car following models available in R for numerical simulation.

## Installation

`carfollowingmodels`is not on CRAN yet. You can download the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("durraniu/carfollowingmodels")
```
## Example

To use any car-following model, you need to provide the lead vehicle data, initial position, speed and/or acceleration of following vehicle(s), and model parameters.  

Following shows an example with 5 following vehicles. The lead vehicle is moving at 13.9 m/s at the reference position of 100 m.

```{r example}
# Time
last_time <- 3000 ## s
time_frame <- 0.1 ## s
Time <- seq(from = 0, to = last_time, by = time_frame)
time_length <- length(Time)



## Lead vehicle
vn1_first <- 13.9 ## first speed m/s
xn1_first <- 100 ## position of lead vehicle front center m
bn1_complete <- c(rep(0, 29500),
                 rep(-5, time_length - 29500))



#############################################
### Complete speed trajectory of Lead vehicle
#############################################

vn1_complete <- rep(NA_real_, time_length) ### an empty vector
xn1_complete <- rep(NA_real_, time_length) ### an empty vector

vn1_complete[1] <- vn1_first
xn1_complete[1] <- xn1_first

for (t in 2:time_length) {

 ### Lead vehicle calculations
 vn1_complete[t] <- vn1_complete[t-1] + (bn1_complete[t-1] * time_frame)

 vn1_complete[t] <- ifelse(vn1_complete[t] < 0, 0, vn1_complete[t])


 xn1_complete[t] <- xn1_complete[t-1] + (vn1_complete[t-1] * time_frame) +
  (0.5 * bn1_complete[t-1] * (time_frame)^2)

}

## Lead vehicle data in a dataframe
ldf <- data.frame(Time, bn1_complete, xn1_complete, vn1_complete)
```


To predict the trajectories of the 5 following vehicles, you can use any car-following model available in this package. For example, the `simulate_idm2()` function uses the Intelligent Driver Model as shown below. For more details on input arguments, type `?simulate_idm2`in the console.

```{r exampleCont}
library(carfollowingmodels)

## Run the IDM function:
results <- simulate_idm2(

resolution=0.1,
N=5,

dfn1=ldf,
xn1="xn1_complete",
vn1="vn1_complete",

xn_first=list(85, 70, 55, 40, 25),
vn_first=list(12, 12, 12, 12, 12),
ln=list(5, 5, 5, 5, 5),

a=2,
v_0=14.4,
small_delta=1,
s_0=4,
Tg=1,
b=1.5
)


head(results)
```


Now you can plot the results:

```{r plots}
library(ggplot2)

## Position
results_at_time_0_LV <- subset(results, fvn==1 & Time ==0)
results_at_time_0_FV <- subset(results, Time ==0)

ggplot() +
  geom_rect(data = results_at_time_0_LV,
            aes(xmin = xn1 - ln1,
                xmax = xn1,
                
                ymin = 0.628,
                ymax = 3.028)) +
  geom_rect(data = results_at_time_0_FV,
            aes(group = fvn,
                fill = as.factor(fvn),
                xmin = xn - 5,
                xmax = xn,
                
                ymin = 0.628,
                ymax = 3.028)) +
  geom_hline(yintercept = 3.6, linetype = "longdash") +
  coord_fixed(ratio=1)


## Speed
ggplot(data = results) +
  geom_line(aes(x = Time, y = vn, color = as.factor(fvn), group=fvn)) +
  geom_line(data = subset(results, fvn==1),
            aes(x = Time, y = vn1, color = "LV Speed")) +
  theme(legend.title = element_blank())
```

